---
title: "Homework 1 - cz2657"
output: github_document
---

## Problem 1

```{r dataframe}

library(tidyverse) 

set.seed(1234)

hw1df_df = tibble (
  x = rnorm(10), 
  vec_logical = c(x>0), 
  vec_char = c("today", "is", "Thursday", "and", "I", "am", "excited", "for", "Friday", "tomorrow"), 
  vec_factor = (factor(c("yes", "no", "maybe", "maybe", "yes", "maybe", "no", "yes", "yes", "no")))
)

mean(pull(hw1df_df, x))
mean(pull(hw1df_df, vec_logical))
mean(pull(hw1df_df, vec_char))
mean(pull(hw1df_df, vec_factor))
```

It appears that we can take the mean of a numeric and logical variables, but not character or factor variables, according to the R warning messages 

`as.numeric(vec_logical)`
`as.numeric(vec_char)`
`as.numeric(vec_factor)`

When attempting to convert logical or character variables to numeric variables, RStudio produces an error that the object `vec_logical` or `vec_char` is not found. However, it will convert a factor variable into numeric levels in alphabetical order. 

This does explain why it is not possible to take the mean of a character variable, as R cannot deal with it numerically. It also likely explains that R cannot take the mean of a factor variable unless it is converted into a numeric variable. Finally, it is not clear why R can take the mean of the logical variable (presumably counting TRUE as 1 and FALSE as 0), but cannot assign those values to the variable in converting to a numeric variable. 

**WHY NO MEAN but, yes convert and vise versa? 

## Problem 2

```{r}
data("penguins", package = "palmerpenguins")

penguins 

mean(pull(penguins, flipper_length_mm), na.rm=TRUE)
```

The `penguins` dataset describes the a group of penguins. Important variables include `species`, `island`, and `sex`, which indicate the type of penguin and where he/she resides. Other variables describe the size of the penguin, including `bill_length_mm`, `bill_depth_mm`, `flipper_length_mm`, and `body_mass_g`. Finally, the variable `year` is included for each penguin. The dataset has `r nrow(penguins)` rows or observations and `r ncol(penguins)` columns or variables. The mean flipper length is `r mean(pull(penguins, flipper_length_mm), na.rm=TRUE)`, excluding NA values. 
