---
title: "Homework 1 - cz2657"
output: github_document
---

## Problem 1

```{r dataframe}

library(tidyverse) 

set.seed(1234)

hw1df_df = tibble (
  x = rnorm(10), 
  vec_logical = c(x>0), 
  vec_char = c("today", "is", "Thursday", "and", "I", "am", "excited", "for", "Friday", "tomorrow"), 
  vec_factor = (factor(c("yes", "no", "maybe", "maybe", "yes", "maybe", "no", "yes", "yes", "no")))
)

mean(pull(hw1df_df, x))
mean(pull(hw1df_df, vec_logical))
mean(pull(hw1df_df, vec_char))
mean(pull(hw1df_df, vec_factor))

```

It appears that we can take the mean of a numeric and logical variables, but not character or factor variables, according to the R warning messages 

Now let's try to convert to numeric: 

`as.numeric(pull(hw1df_df, vec_logical))`
`as.numeric(pull(hw1df_df, vec_char))`
`as.numeric(pull(hw1df_df, vec_factor))`

We can convert logical and factor variables to numeric values, but RStudio produces an error when trying to convert character variables to numeric values. 

This explains why it is not possible to take the mean of a character variable, as R cannot deal with it numerically. 

It also demonstrates that R cannot take the mean of a factor variable unless it is converted into a numeric variable. 

Finally, RStudio must store "TRUE" and "FALSE" as 1 and 0, respectively, for logical variables, so it assigns those values explicitly when taking the mean but shows "TRUE" and "FALSE" in the dataframe. 

## Problem 2

```{r}
data("penguins", package = "palmerpenguins")

penguins 

```

The `penguins` dataset describes the a group of penguins. Important variables include `species`, `island`, and `sex`, which indicate the type of penguin and where he/she resides. Other variables describe the size of the penguin, including `bill_length_mm`, `bill_depth_mm`, `flipper_length_mm`, and `body_mass_g`, which provide specific measurements of each penguin. Finally, the variable `year` is included for each penguin. The dataset has `r nrow(penguins)` rows or observations and `r ncol(penguins)` columns or variables. The mean flipper length is `r mean(pull(penguins, flipper_length_mm), na.rm=TRUE)`, excluding NA values. 

Now, I'll construct a plot

```{r}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species))+geom_point()

ggsave("first_scatterplot.pdf")
```

The plot is saved as a pdf. 

