---
title: "Homework 1 - cz2657"
output: github_document
---

## Problem 1

```{r, include = FALSE}

library(tidyverse) 

```

```{r}
set.seed(1234)

hw1df_df = tibble (
  x = rnorm(10), 
  vec_logical = c(x>0), 
  vec_char = c("today", "is", "Thursday", "and", "I", "am", "excited", "for", "Friday", "tomorrow"), 
  vec_factor = (factor(c("yes", "no", "maybe", "maybe", "yes", "maybe", "no", "yes", "yes", "no")))
)

mean(pull(hw1df_df, x))
mean(pull(hw1df_df, vec_logical))
mean(pull(hw1df_df, vec_char))
mean(pull(hw1df_df, vec_factor))
```


We can take the mean of a numeric and logical variables, but not character or factor variables.  

Now let's try to convert the variables to numeric: 

```{r, results = 'hide', warning = FALSE}
as.numeric(pull(hw1df_df, vec_logical))

as.numeric(pull(hw1df_df, vec_char))

as.numeric(pull(hw1df_df, vec_factor))
```

We can convert logical and factor variables to numeric values, but RStudio produces an error when trying to convert character variables to numeric values.

*   This explains why it is not possible to take the mean of a character variable, as R cannot deal with it numerically. 

*   It also demonstrates why R cannot take the mean of a factor variable, as it must first be converted into a numeric variable. 

*   For logical variables, RStudio seems to store "TRUE" and "FALSE" as 1 and 0, respectively, so it assigns those values explicitly when taking the mean, but shows only "TRUE" and "FALSE" in the dataframe. Therefore, a logical variable can be readily converted into a numeric variable. 

## Problem 2

```{r}
data("penguins", package = "palmerpenguins")
```

The `penguins` dataset describes the a group of penguins. Important variables include `species`, `island`, and `sex`, which indicate the type of penguin and where he/she resides. Other variables describe the size of the penguin, including `bill_length_mm`, `bill_depth_mm`, `flipper_length_mm`, and `body_mass_g`. Finally, the variable `year` is included for each penguin, which spans the years `r min(pull(penguins, year))` to `r max(pull(penguins, year))`. The dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns. The mean flipper length is `r mean(pull(penguins, flipper_length_mm), na.rm=TRUE)` mm, excluding missing values. 

Now, I'll construct a plot

```{r, warning = FALSE, message = FALSE}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species))+geom_point()

ggsave("first_scatterplot.pdf")
```

The plot is saved as a pdf. 

